### 2020-07-26 inst/sample_size_estimator/www/example.Rmd
added shiny app to package with rank and protein selection

<html>
<div class = "row">
<div class = "col-md-6">
<h3>Protein Abundance File</h3>
<p>This must be a csv file in the <b>wide format</b>. Rows are proteins, columns are 
    Biological replicates or samples, cells are protein  abundances. The following
    image is an example, example data file can be found <a href = "https://github.com/Vitek-Lab/MSstatsSampleSize-Shiny/blob/master/OV_SRM_train.csv " target="_blank"><b>here</b></a>
    </p>
<img src = ab.PNG width = "100%" height = "350">

</div>

<div class = "col-md-6">
<h3>Sample Annotation File</h3>
<p> This must be a csv file in the <b>long format</b>. It includes columns BioReplicate
    for sample ID and Condition for group information. BioReplicate information
    should be the same as the column names of Preliminary data. The following
    image is an example, example data file can be found <a href = "https://github.com/Vitek-Lab/MSstatsSampleSize-Shiny/blob/master/OV_SRM_train_annotation.csv" target="_blank"><b>here</b></a>
    </p>
<img src = an.PNG  height = "350">
</div>
</div>
</html>

### 2020-07-26 R/run_estimator.R

@@ -1,0 +1,9 @@
+#' @export
+run_estimator <- function() {
+  appDir <- system.file("sample_size_estimator", package = "MSstatsSampleSize")
+  if (appDir == "") {
+    stop("Could not find example directory. Try re-installing `MSstatsSampleSize`.", call. = FALSE)
+  }
+  
+  shiny::runApp(appDir, display.mode = "normal")
+}

### 2020-07-26 inst/sample_size_estimator/global.R

@@ -1,0 +1,156 @@
+suppressMessages({
+  library(shiny)
+  library(shinydashboard)
+  library(shinyWidgets)
+  library(shinycssloaders)
+  library(ggplot2)
+  library(data.table)
+  library(dplyr)
+  library(MSstatsSampleSize)
+  #### Caret package dependencies ####
+  library(e1071)
+  library(kernlab)
+  library(naivebayes)
+  library(randomForest)
+})
+
+wd <- getwd()
+src_dir <- file.path(wd,"scripts")
+req_files <- list.files(src_dir, full.names = T)
+  
+for (i in req_files) {
+  tryCatch({
+    source(i)
+    message(sprintf("Sourced file : %s", i))
+  }, error = function(e){
+    print(conditionMessage(e))
+    print(sprintf("Error in sourcing : %s", i))
+  })
+  
+}
+conn <<- .logGeneration()
+message("\n Log File saved at", conn$file)
+
+#### GLOBAL VARS ####
+FORMATS_LIST <- list("Protein-level quantification" = "standard", 
+                     "Example from MSstatsSampleSize" = "examples")
+FORMATS <- c("examples", "standard")
+EXTENSTIONS <- c("text/csv",
+                 "text/comma-separated-values,text/plain",
+                 ".csv", "text/tab-separated-values", ".tsv")
+
+MODELS <- c('rf','nnet','svmLinear','logreg','naive_bayes')
+names(MODELS) <- c("Random Forest", "Neural Network",
+                   "Support Vector Machines with Linear Kernel",
+                   "Logistic Regression", "Naive Bayes")
+
+STOPPING_METRIC <- c("AUTO", "deviance", "logloss", "MSE", "RMSE", "MAE", "RMSLE",
+                     "AUC", "lift_top_group", "misclassification", "AUCPR",
+                     "mean_per_class_error")
+
+FOLD_ASSIGNMENT <- c("AUTO", "Random", "Modulo", "Stratified")
+
+
+FAMILY <-  c("gaussian", "binomial", "quasibinomial", "ordinal", "multinomial",
+           "poisson", "gamma", "tweedie", "negativebinomial")
+
+SOLVER <- c("AUTO", "IRLSM", "L_BFGS", "COORDINATE_DESCENT_NAIVE", 
+            "COORDINATE_DESCENT", "GRADIENT_DESCENT_LH", "GRADIENT_DESCENT_SQERR")
+
[…]

### 2020-07-26 inst/sample_size_estimator/scripts/Methods.R

@@ -541,7 +541,7 @@ theme_MSstats <- function(x.axis.size = 10, y.axis.size = 10,
         geom_line(aes(y = mean_acc, group = 1), size = 0.75, color = "blue")+
         labs(x = "Simulated Sample Size Per Group", y = "Predictive Accuracy",
              #title = sprintf("Classifier %s", alg),
-             subtitle = sprintf("Optimum accuracy achieved when sample size per group is : %s",
+             subtitle = sprintf("Optimal sample size per group is : %s",
                                 optimal_ss))+
         scale_y_continuous(breaks = scales::pretty_breaks(), limit = y_lim)+
         scale_x_continuous(breaks = unique(df$sample))+
@@ -667,8 +667,7 @@ qc_boxplot <- function(data = NULL ,annot = NULL){
 #' @param k A integer value for the number of features to select
 #' @param session A session object for the shiny app
 #' @keywords internal
-ss_classify_caret <- function(n_samp, sim_data, classifier, k = 10,
-                              session = NULL, ...){
+ss_classify_caret <- function(n_samp, sim_data, classifier, k = 10, ...){
     
     dots <- list(...)
     session <- dots$session
@@ -749,7 +748,7 @@ simulate_grid <- function(data = NULL, annot = NULL, num_simulation, exp_fc,
                     shiny::need(all(samp >= 1), "All samples Need to be >= 1"))
     
     
-    if(rank_proteins == "combined"){
+    if(rank_proteins == "Combined"){
         protein_quantile_cutoff <- list(Mean = mean_ip, SD = sd_ip)
         protein_select <- list(Mean = mean_equality, SD = sd_equality)
     }else if(rank_proteins == "Mean"){
@@ -778,14 +777,14 @@ simulate_grid <- function(data = NULL, annot = NULL, num_simulation, exp_fc,
         sim[[paste(i)]] <- simulateDataset(data = data,
                                            annotation = annot,
                                            num_simulations = num_simulation,
-                                           protein_select = protien_select,
+                                           protein_select = protein_select,
                                            protein_quantile_cutoff = protein_quantile_cutoff, 
                                            expected_FC = fc,
                                            list_diff_proteins =  diff_prots,
                                            samples_per_group = i,
                                            simulate_validation = as.logical(sim_valid),
                                            valid_samples_per_group = valid_samples_per_grp,
-                                           session = session, ...)
+                                           session = session)
     }
[…]

### 2020-07-26 inst/sample_size_estimator/server.R

@@ -1,0 +1,610 @@
+
+function(session, input, output) {
+  # enables the helper functionality in the UI
+  shinyhelper::observe_helpers(help_dir = "help_mds")
+  
+  # stop the h2o cluster once the app is shutdown
+  onStop(function() {
+    try({
+      h2o::h2o.shutdown(prompt = F)
+      }, silent = T)
+    .status("Shutting Down Application", conn = conn$con)
+    
+    if(isOpen(conn$con)){
+      close(conn$con)
+      rm(list = ls())
+    }
+    stopApp()
+  })
+  
+  # List of reactive values
+  rv <- reactiveValues()
+  rv$seed <- -1
+  rv$use_h2o <- F
+  
+  observeEvent(input$start_process,{
+    updateTabsetPanel(session = session, inputId = "tabs", "import_data")
+  })
+  ## Set maximum size of uploaded files to 300mb
+  options(shiny.maxRequestSize = 300*1024^2)
+  #### Toggle control for sidebar ####
+  # Enable or disable fileInputs based on type of data selected
+  observeEvent(input$data_format,{
+    ids = c("standard_count", "standard_annot")
+    lapply(ids, shinyjs::toggleElement, anim = T, animType = "fade",
+           condition = input$data_format == "standard")
+  })
+
+  #### Import data, action click ####
+  data <- eventReactive(input$import_data, {
+    SIM_CHOICES <<-0
+    rv$seed <- -1
+    rv$use_h2o <- F
+    rv$classification <- NULL
+    
+    disable_btns <- c("download_plots", "back_varimp", "fwd_varimp", 
+                      "generate_report", "run_model","nav_to_exp",
+                      "fwd","back","download_pca", "run_model")
+    
+    lapply(disable_btns, shinyjs::disable)
+    updateSelectInput(session = session, inputId ="simulations", label = "Simulations",
+                      choices = SIM_CHOICES)
+    output$default <- renderText("")
+
+    withProgress({
+      data <- .catch_faults(
+        expr = format_data(format = input$data_format, count = input$standard_count,
+                           annot = input$standard_annot, session = session,
+                           conn = conn),
+        session = session, conn = conn
+      )
+      
+      # Create a global variable for dynamic ui element
+      B_GROUP <<- unique(data$annot_data$Condition) 
+      # update the slider input with the maximum number of proteins found in the dataset
+      updateSliderInput(session = session, inputId = "prot_num",
+                        min = 1, max = data$n_prot, value = data$n_prot)
[…]

### 2020-07-26 inst/sample_size_estimator/ui.R

@@ -1,0 +1,422 @@
+## MSstatsSampleSize Shiny UI
+###############################################################
+dashboardPage(
+  skin="black",
+  tags$head(
+    tags$style(
+      HTML(".shiny-notification {
+              height: 100px;
+              width: 800px;
+              opacity: 1;
+              position:fixed;
+              top: calc(50% - 50px);;
+              left: calc(55% - 400px);;
+            }
+           .shiny-output-error-validation {
+           color: red;
+           }"
+      )
+    )
+  ),
+  ##### Header #####
+  header =  dashboardHeader(
+    title = "MSstats- Sample Size Estimator",
+    titleWidth = 350
+  ),
+  #### SideBar ####
+  sidebar =  dashboardSidebar(
+    shinyjs::useShinyjs(),
+    width = 250,
+    sidebarMenu(
+      id="tabs",
+      ##### Home Tab ####
+      menuItem("Home", tabName = "home", icon = icon("home")),
+      #### Data Import Tab ####
+      menuItem("1: Import data", tabName = "import_data", icon = icon("file-import")),
+      #### Data Simulation Tab ####
+      menuItem("2: Simulate datasets", 
+               tabName = "explore_simulated", icon = icon("project-diagram")),
+      #### Analyze Simulation Tab ####
+      menuItem("3: Plan experiment", 
+               tabName = "plan_experiment", icon = icon("vial"))
+    )
+  ),
+  ##### Body ####
+  body = dashboardBody(
+    shinyjs::useShinyjs(),
+    tabItems(
+      #### Home Tab ####
+      tabItem(
+        tabName = "home",
+        includeMarkdown("www/Welcome.md"),
+        actionButton(inputId = "start_process",
+                     label = "Start",
+                     icon = icon("arrow-right"),
+                     style = CSS_BUTTON,            
+                     width = '200px')
+      ),
+      #### Import Data Tab ####
+      tabItem(
+        tabName = "import_data",
+        navbarPage("", id = "myNavBar",
+                   tabPanel("Import", fluidPage( 
+                     box(title = "Data Import Wizard", width = 12,
+                         solidHeader = T, status = 'primary',
+                         includeHTML("www/data_import.Rmd"),
+                         fluidRow(
+                           column(width = 4, 
+                                  selectInput("data_format", "Select Data Type",
+                                              choice = FORMATS_LIST,
+                                              selected = FORMATS_LIST[2])),
+                           column(width = 4, 
[…]

### 2020-07-27 inst/sample_size_estimator/scripts/Methods.R
animation for mean and sd selections

@@ -696,20 +696,6 @@ ss_classify_caret <- function(n_samp, sim_data, classifier, k = 10, ...){
 #' @title Simulate datasets to be tested out
 #' @description A wrapper function for the `simulateDataset` function from the 
 #' MSstatsSampleSize package which enables simulating datasets for running experiments
-#' @param data 
-#' @param annot
-#' @param num_simulations
-#' @param exp_fc
-#' @param list_diff_proteins
-#' @param sel_simulated_proteins
-#' @param prot_proportion
-#' @param prot_number
-#' @param samples_per_group
-#' @param sim_valid
-#' @param valid_samples_per_grp
-#' @param seed
-#' @param session
-#' @return 
 simulate_grid <- function(data = NULL, annot = NULL, num_simulation, exp_fc,
                           list_diff_proteins, samples_per_group, sim_valid,
                           rank_proteins, mean_ip, mean_equality, sd_ip,
@@ -791,6 +777,7 @@ simulate_grid <- function(data = NULL, annot = NULL, num_simulation, exp_fc,
     return(sim)
 }
 
+
 .protein_select<- function(df, cutoff = list("Mean" = 0.2, "SD" = 0.2),
                           equality = list("Mean" = "high", "SD" = "high"),
                           conn){
@@ -808,6 +795,11 @@ simulate_grid <- function(data = NULL, annot = NULL, num_simulation, exp_fc,
 }
 
 
+#' @title Evaluate quantiles given vectors and probability
+#' @param df A data.frame, with column names for which quantiles are calculated
+#' @param vec_name A vector with the name of the column
+#' @param prob A numeric value for the probablitiy
+#' @param rnd A integer value which defines where to round the quantiles
 .quantile_eval <- function(df, vec_name, prob, eq, rnd = 4){
     eq_vals <- c(">","<=")
     names(eq_vals) <- c("high", "low")

### 2020-07-27 inst/sample_size_estimator/server.R

@@ -240,6 +240,22 @@ function(session, input, output) {
                            condition = input$upload_params != T)
   })
   
+  observeEvent(input$rank_proteins,{
+    s_ids <- c('sd_ip', 'sd_equality')
+    m_ids <- c('mean_ip', "mean_equality")
+    if(input$rank_proteins == "Mean"){
+      lapply(s_ids, shinyjs::hide, anim = TRUE)
+      lapply(m_ids, shinyjs::show, anim = TRUE)
+    }else if(input$rank_proteins == "SD"){
+      lapply(m_ids, shinyjs::hide, anim = TRUE)
+      lapply(s_ids, shinyjs::show, anim = TRUE)
+    }else{
+      ids <- c(s_ids, m_ids)
+      lapply(ids, shinyjs::show, anim = TRUE)
+    }
+  })
+  
+  
   #### Simulate Data Button Click ####
   simulations <- eventReactive(input$simulate,{
     withProgress({
@@ -254,7 +270,6 @@ function(session, input, output) {
       if(input$set_seed){
         rv$seed <- input$seed
       }
-      
       data <- simulate_grid(data = data()$wide_data,
                             annot = data()$annot_data,
                             est_var = data()$est_var,
@@ -309,15 +324,30 @@ function(session, input, output) {
     s2 <- ifelse(input$sd_equality == "High",
                  ggplot_build(mean_sd_plt())$layout$panel_params[[1]]$y.range[2],
                  0)
+    if(input$rank_proteins == 'Mean'){
+      s1 <- s2 <- 0
+    }else if(input$rank_proteins == 'SD'){
+      m1 <- m2 <- 0
+    }
     
     output$mean_sd_plot <- renderPlot({
       mean_sd_plt()+
-        geom_hline(yintercept = s1, linetype = 'dashed', color = 'red')+
-        geom_vline(xintercept = m1, linetype = 'dashed', color = 'green')+
+        geom_hline(yintercept = s1, linetype = 'dashed', color = 'green')+
+        geom_vline(xintercept = m1, linetype = 'dashed', color = 'red')+
         annotate('rect', xmin = ifelse(m1>m2,m2,m1), xmax = ifelse(m1>m2,m1,m2),
                  ymin = ifelse(s1>s2,s2,s1), ymax = ifelse(s1>s2,s1,s2),
                  alpha = 0.1, fill = 'yellow')
     })
+    
+    # output$adv_meansd <- renderText({
[…]

### 2020-07-27 inst/sample_size_estimator/ui.R

@@ -251,35 +251,38 @@ dashboardPage(
                  box(title = "Mean-Variance Plot", solidHeader = T,
                      status = "primary", width = 12,
                      withSpinner(plotOutput("mean_sd_plot"),
-                                 proxy.height = "200px")
-                 ),
+                                 proxy.height = "200px"),
+                     htmlOutput("adv_meansd")
                  )
+          )
         ),
-        #### Simulation details section #####
-        box(title = "Simulated Datasets", solidHeader = T, status = "primary",
-            width = 12,
-            fluidRow(
-              ##### Select individual simulation, previous and next buttons #####
-              column(4, selectInput(inputId = "simulations", label = "Simulations", 
-                                    choices = NULL)),
-              column(2, shinyjs::disabled(actionButton(inputId = "back",
-                                                       label = "Previous",
-                                                       icon = icon("arrow-left"),
-                                                       style = "margin-top: 25px;",
-                                                       width = '100px'))),
-              column(2, shinyjs::disabled(actionButton(inputId = "fwd",
-                                                       label = "Next",
-                                                       icon = icon("arrow-right"),
-                                                       style = "margin-top: 25px;",
-                                                       width = '100px'))),
-              column(2, shinyjs::disabled(downloadButton(outputId = "download_pca",
-                                                         label = "Download",
-                                                         icon = icon("download"),
+        fluidRow(
+          #### Simulation details section #####
+          box(title = "Simulated Datasets", solidHeader = T, status = "primary",
+              width = 12,
+              fluidRow(
+                ##### Select individual simulation, previous and next buttons #####
+                column(4, selectInput(inputId = "simulations", label = "Simulations", 
+                                      choices = NULL)),
+                column(2, shinyjs::disabled(actionButton(inputId = "back",
+                                                         label = "Previous",
+                                                         icon = icon("arrow-left"),
                                                          style = "margin-top: 25px;",
-                                                         width = '150px')))
-              
-            ),
-            withSpinner(plotOutput("pca_plot", height = '600px'))
+                                                         width = '100px'))),
+                column(2, shinyjs::disabled(actionButton(inputId = "fwd",
+                                                         label = "Next",
+                                                         icon = icon("arrow-right"),
+                                                         style = "margin-top: 25px;",
+                                                         width = '100px'))),
+                column(2, shinyjs::disabled(downloadButton(outputId = "download_pca",
+                                                           label = "Download",
+                                                           icon = icon("download"),
[…]

### 2020-07-29 inst/sample_size_estimator/scripts/Methods.R
file upload for DA proteins

@@ -544,7 +544,8 @@ theme_MSstats <- function(x.axis.size = 10, y.axis.size = 10,
              subtitle = sprintf("Optimal sample size per group is : %s",
                                 optimal_ss))+
         scale_y_continuous(breaks = scales::pretty_breaks(), limit = y_lim)+
-        scale_x_continuous(breaks = unique(df$sample))+
+        scale_x_continuous(breaks = df$sample,
+                           labels = paste(df$sample))+
         theme_MSstats(...)+
         theme(plot.subtitle = element_text(face = "italic", color = "red"))
     
@@ -710,11 +711,6 @@ simulate_grid <- function(data = NULL, annot = NULL, num_simulation, exp_fc,
     if(exp_fc != "data"){
         .status(detail = "Extracting Fold Change Informations", value = 0.15, 
                session = session, log = conn$con)
-        .status(detail = sprintf("List of differential proteins selected: (%s)", 
-                                list_diff_proteins), session = session, 
-                log = conn$con)
-        
-        diff_prots <- unlist(strsplit(list_diff_proteins, ","))
         fc <- exp_fc$`Fold Change Value`
         names(fc) <- exp_fc$orig_group
     } else{
@@ -766,7 +762,7 @@ simulate_grid <- function(data = NULL, annot = NULL, num_simulation, exp_fc,
                                            protein_select = protein_select,
                                            protein_quantile_cutoff = protein_quantile_cutoff, 
                                            expected_FC = fc,
-                                           list_diff_proteins =  diff_prots,
+                                           list_diff_proteins =  list_diff_proteins,
                                            samples_per_group = i,
                                            simulate_validation = as.logical(sim_valid),
                                            valid_samples_per_group = valid_samples_per_grp,

### 2020-07-29 inst/sample_size_estimator/scripts/simulateDataset.R

@@ -285,7 +285,7 @@ simulateDataset <- function(data, annotation, log2Trans = FALSE,
             baseline <- names(expected_FC)[expected_FC == 1] # baseline group
             otherlines <- names(expected_FC)[expected_FC != 1] # compared groups
             
-            if(!all(list_diff_proteins %in%  selectedPros)){
+            if(all(list_diff_proteins %in%  selectedPros)){
                 selected_diff_proteins <- intersect(list_diff_proteins, selectedPros)
                 
                 if (length(selected_diff_proteins) == 0) {

### 2020-07-29 inst/sample_size_estimator/server.R

@@ -37,7 +37,6 @@ function(session, input, output) {
 
   #### Import data, action click ####
   data <- eventReactive(input$import_data, {
-    SIM_CHOICES <<-0
     rv$seed <- -1
     rv$use_h2o <- F
     rv$classification <- NULL
@@ -47,10 +46,9 @@ function(session, input, output) {
                       "fwd","back","download_pca", "run_model")
     
     lapply(disable_btns, shinyjs::disable)
-    updateSelectInput(session = session, inputId ="simulations", label = "Simulations",
-                      choices = SIM_CHOICES)
-    output$default <- renderText("")
-
+    updateSelectInput(session = session, inputId ="simulations",
+                      label = "Simulations", choices = SIM_CHOICES)
+    
     withProgress({
       data <- .catch_faults(
         expr = format_data(format = input$data_format, count = input$standard_count,
@@ -127,16 +125,15 @@ function(session, input, output) {
     
     lapply(ids, shinyjs::toggleElement, anim = T, animType = "fade",
            condition = input$exp_fc != T)
-    
+  })
+  
+  observeEvent(input$da_prots,{
     ids <- c("da_prots_file", "diff_prot", "diff_prot_help")
-    cond <- c(input$da_prots == T && input$exp_fc != T, 
-              input$da_prots == F && input$exp_fc != T)
-    
-    if(cond[1]){
+    if(input$da_prots == "Upload csv file" && input$exp_fc != T){
       shinyjs::hideElement(id = "diff_prot", anim = T, animType = "fade")
       shinyjs::hideElement(id = "diff_prot_help", anim = T, animType = "fade")
       shinyjs::showElement(id = "da_prots_file", anim = T, animType = "fade")
-    } else if(cond[2]){
+    } else if(input$da_prots == "Text Input" && input$exp_fc != T){
       shinyjs::hideElement(id = "da_prots_file", anim = T, animType = "fade")
       shinyjs::showElement(id = "diff_prot", anim = T, animType = "fade")
       shinyjs::showElement(id = "diff_prot_help", anim = T, animType = "fade")
@@ -145,35 +142,10 @@ function(session, input, output) {
     }
   })
   
-  observeEvent(input$da_prots,{
-    ids <- c("da_prots_file", "diff_prot", "diff_prot_help")
-    if(length(input$da_prots) >1 || is.null(input$da_prots)){
-      lapply(ids, shinyjs::hideElement, anim = T, animType = "fade")
-      
[…]

### 2020-07-29 inst/sample_size_estimator/ui.R

@@ -186,11 +186,10 @@ dashboardPage(
                                     id = "fc_values_help"),
               br(),
               # Input vector off different proteins
-              
-              checkboxGroupInput(inputId = "da_prots", label = "How do you want to specify Differentially
-                                 abundant proteins ?", inline = T,
-                                 choiceNames = list("Text input", "Upload csv file"),
-                                 choiceValues = list(F, T), selected = F),
+              selectInput(inputId = "da_prots",
+                          label = paste("How do you want to specify Differentially",
+                                        "abundant proteins ?"),
+                          choices = c("Text Input", "Upload csv file")),
               
               textInput(inputId = "diff_prot", 
                         label = "List of Differential Abundant Proteins",

### 2020-07-30 inst/sample_size_estimator/scripts/Methods.R
debug pca and accuracy plots

@@ -534,18 +534,15 @@ theme_MSstats <- function(x.axis.size = 10, y.axis.size = 10,
 #' @import ggplot2
 #' @importFrom scales pretty_breaks
 .plot_acc <- function(df, y_lim, optimal_ss, ...){
-    g <- ggplot(data = df, aes(x = sample))+
+    g <- ggplot(data = df, aes(x = as.factor(sample)))+
         geom_boxplot(aes(y = acc, group = sample, fill = fill_col), alpha = 0.5)+
         scale_fill_identity()+
         geom_point(aes(y = mean_acc))+
         geom_line(aes(y = mean_acc, group = 1), size = 0.75, color = "blue")+
         labs(x = "Simulated Sample Size Per Group", y = "Predictive Accuracy",
-             #title = sprintf("Classifier %s", alg),
              subtitle = sprintf("Optimal sample size per group is : %s",
                                 optimal_ss))+
         scale_y_continuous(breaks = scales::pretty_breaks(), limit = y_lim)+
-        scale_x_continuous(breaks = df$sample,
-                           labels = paste(df$sample))+
         theme_MSstats(...)+
         theme(plot.subtitle = element_text(face = "italic", color = "red"))
     

### 2020-07-30 inst/sample_size_estimator/server.R

@@ -127,7 +127,7 @@ function(session, input, output) {
            condition = input$exp_fc != T)
   })
   
-  observeEvent(input$da_prots,{
+  observe({
     ids <- c("da_prots_file", "diff_prot", "diff_prot_help")
     if(input$da_prots == "Upload csv file" && input$exp_fc != T){
       shinyjs::hideElement(id = "diff_prot", anim = T, animType = "fade")
@@ -225,7 +225,6 @@ function(session, input, output) {
   #### Simulate Data Button Click ####
   simulations <- eventReactive(input$simulate,{
     withProgress({
-      browser()
       exp_fc <- ifelse(input$exp_fc, 'data', '')
       if(exp_fc ==""){
         baseline <- data.table(Group = input$b_group,
@@ -351,18 +350,23 @@ function(session, input, output) {
       withProgress({
         p <- list()
         library(gridExtra)
-        pdf(file = file)
+        pdf(file = file, height = 11, width = 8.5)
         for(i in seq_along(SIM_CHOICES)){
           .status(sprintf("Plotting %s plot", i), 
                  value = i/length(SIM_CHOICES), session = session)
           vals <- unlist(stringr::str_extract_all(SIM_CHOICES[i],'\\d+'))
           sim <- sprintf("simulation%s",vals[1])
-          p <- append(p, list(make_pca_plots(simulations()[[vals[2]]], 
-                                             which = sim, dot_size=1,
-                                             title = SIM_CHOICES[i],
-                                             x.axis.size = 4, y.axis.size = 4,
-                                             margin = 0.2, legend.size = 7,
-                                             download = T)
+          p <- append(p, list(designSampleSizePCAplot(simulations()[[vals[2]]], 
+                                             which = sim, conn= conn,
+                                             session = session, save.pdf = F,
+                                             dot_size = 2)+
+                                theme(axis.text.x = element_text(size = 5, colour = "black"), 
+                                      axis.text.y = element_text(size = 5, colour = "black"),
+                                      axis.ticks = element_line(colour = "black"), 
+                                      axis.title.x = element_text(size = 9, vjust = -0.4), 
+                                      axis.title.y = element_text(size = 9, vjust = 0.3),
+                                      title = element_text(size = 8, vjust = 1.5),
+                                      plot.margin = unit(rep(0.5, 4), "cm"))
                               )
                       )
           

### 2020-07-30 inst/sample_size_estimator/Report.Rmd
remove offensive error message, updated report for new params

---
title: "MSstats Sample Size Report"
author: "MSstatsSampleSize GUI"
output: pdf_document
params:
  data: NA
  use_h2o: NA
  alg: NA
  sample: NA
  annot: NA
  count: NA
  n_sim: NA
  fc: NA
  list_diff_prots: NA
  fc_values: NA
  proportion: NA
  number: NA
  valid: NA
  valid_sample: NA
  sim_by: NA
  set_seed: NA
  seed: NA
  baseline: NA

---

```{r wrap-hook, include=FALSE, eval=TRUE}
library(knitr)
hook_output = knit_hooks$get('output')
knit_hooks$set(output = function(x, options) {
  # this hook is used only when the linewidth option is not NULL
  if (!is.null(n <- options$linewidth)) {
    x = knitr:::split_lines(x)
    # any lines wider than n should be wrapped
    if (any(nchar(x) > n)) x = strwrap(x, width = n)
    x = paste(x, collapse = '\n')
  }
  hook_output(x, options)
})
```

# Result

```{r echo=FALSE, warning=FALSE, fig.width= 8.5, fig.height=8.5, message=FALSE}
p1 <- designSampleSizeClassificationPlots(data = params$data,
                                          protein_importance_plot = F,
                                          predictive_accuracy_plot = T,
                                          use_h2o = params$use_h2o,
                                          alg =  params$alg,
                                          save.pdf = F)

p2 <- designSampleSizeClassificationPlots(data = params$data, 
                                        protein_importance_plot = T,
                                        predictive_accuracy_plot = F,
                                        save.pdf = F, session = "")

gridExtra::grid.arrange(p1$plot, p2$plot, ncol = 1)
```

\newpage
# Parameters used to simulate selected sample size
```{r echo=FALSE, linewidth=60}
cat("Input Abundance File path:", params$count)
cat("Input Annotation File path:", params$annot)
cat("Set Seed ?", params$set_seed)
if(params$set_seed){
  cat("Seed Value:", params$seed)
}
cat("Number of Simulations:", params$n_sim)
cat("Expected Fold Change:", !params$fc)
if(!params$fc){
  cat("List of Different Proteins:", params$list_diff_prots)
  cat("Fold Change values:", params$fc_values)
}
if(params$set_seed){
  cat("Seed Value Selected:", params$seed)
}

cat("Samples per group:", gsub("Sample","",params$sample))
cat("Simulate Validation Set:", params$valid)
if(params$valid){
  cat("Validations samples per group:", params$valid_sample)
}
cat("Classifier used to estimate sample size:", params$alg)

```

\newpage
# Code to re-run experiment

```{r eval=FALSE, include=T}
#Need to install MsstatsSampleSize library from Bioconductor if it is not 
#installed

data <-

library(MSstatsSampleSize)

sim_data <- simulateDataset(data,
                            annotation,
                            num_simulations = 10,
                            expected_FC = "data",
                            list_diff_proteins = NULL,
                            select_simulated_proteins = "proportion",
                            protein_proportion = 1,
                            protein_number = 1000,
                            samples_per_group = 50,
                            simulate_validation = FALSE,
                            valid_samples_per_group = 50)

model <- designSampleSizeClassification(simulations = sim_data,
                                        classifier = params$alg)

designSampleSizeClassificationPlots(data = model)

```

\newpage
# Session Information
```{r echo=F}
sessionInfo()
```

### 2020-07-30 inst/sample_size_estimator/scripts/h2o_methods.R

@@ -59,6 +59,7 @@ ss_classify_h2o <- function(n_samp, sim_data, classifier, stopping_metric = "AUT
                             alpha = 0, family, solver, link, min_sdev, laplace, eps,
                             session = NULL){
   
+  config <- h2o_config()
   samp <- unlist(strsplit(n_samp,","))
   config <- h2o_config()
   h2o::h2o.init(nthreads = config$threads, max_mem_size = config$max_mem,
@@ -177,7 +178,6 @@ ss_classify_h2o <- function(n_samp, sim_data, classifier, stopping_metric = "AUT
   }
   rm(valid, valid_x, valid_y, train_x_list, train_y_list)
   gc()
-  h2o:::.h2o.garbageCollect()
   h2o::h2o.shutdown(prompt = F)
   
   return(list("models" = modelz))
@@ -245,17 +245,3 @@ ss_classify_h2o <- function(n_samp, sim_data, classifier, stopping_metric = "AUT
   
   return(output)
 }
-# res3 <- list()
-# 
-# 
-# for(i in seq_along(sim1$simulation_train_Xs)){
-# 
-#   res3[[i]] <- .classification_performance_h2o(index = i, classifier = 'naive_bayes',
-#                                          train_x_list = sim1$simulation_train_Xs,
-#                                          train_y_list = sim1$simulation_train_Ys,
-#                                          valid_x = sim1$valid_X, valid_y = sim1$valid_Y,
-#                                          top_K = 10)
-# 
-# }
-# 
-# 

### 2020-07-30 inst/sample_size_estimator/server.R

@@ -448,7 +448,7 @@ function(session, input, output) {
                                                k = 10, session = session,
                                                conn = conn)
       }else{
-        stop("FUCK OFF!")
+        stop("No h2o for you yet! Go Fish!")
       }
       
     et <- Sys.time()

### 2020-08-14 inst/sample_size_estimator/Report.Rmd
updated ui for data transformation

@@
   fc_values: NA
-  proportion: NA
-  number: NA
   valid: NA
@@
   baseline: NA
-
+  transform: NA
+  rank: NA
+  mean_qq: NA
+  sd_qq : NA
+  mean_eq: NA
+  sd_eq: NA
 ---
@@
 cat("Input Annotation File path:", params$annot)
+cat("Data Transform type: ", params$transform)
 cat("Set Seed ?", params$set_seed)
@@
 if(!params$fc){
+  cat("Basline group:", params$baseline)
   cat("List of Different Proteins:", params$list_diff_prots)
@@
 }
-if(params$set_seed){
-  cat("Seed Value Selected:", params$seed)
+cat("Rank proteins by: ", params$rank)
+if(params$rank %in% c("Mean", "Combined")){
+  cat("Quantile cutoff for Proteins Abundance: ", params$mean_qq, "%\n")
+  cat("Mean Abundance Region: ", params$mean_eq)
 }
-
+if(params$rank %in% c("Combined", "SD")){
+  cat("Quantile cutoff for Standard deviation: ", params$sd_qq, "%\n")
+  cat("Standard DeviationRegion: ", params$sd_eq)
+}
 cat("Samples per group:", gsub("Sample","",params$sample))
@@
 #installed
-
-data <-
 

### 2020-08-14 R/Methods.R

@@ -328,18 +328,19 @@
     consistent_cols <- !setequal(required.annotation, colnames(annotation))
     if(consistent_cols){
         packageStartupMessage(" Failure")
-        nf <- required.annotation[!colnames(annotation) %in% required.annotation]
+        nf <- required.annotation[!required.annotation %in% colnames(annotation)]
         stop("CALL_", func,"_",nf,
              " is not provided in Annotation, please check annotation file")
     }
+    anno_cols <- colnames(data)[!colnames(data) %in% 'Protein']
 
-    ic <- setequal(annotation$BioReplicate, colnames(data)) && 
-        nrow(annotation) == ncol(data)
+    ic <- setequal(annotation$BioReplicate, anno_cols) && 
+        nrow(annotation) == length(anno_cols)
     if(!ic){
         packageStartupMessage(" Failure")
         stop("CALL_",func,"_",
              "Please check the annotation file. 'Bioreplicate' must match with",
-             "the column names of 'data'.")
+             " the column names of 'data'.")
     }
 
     
@@ -534,17 +535,15 @@ theme_MSstats <- function(x.axis.size = 10, y.axis.size = 10,
 #' @import ggplot2
 #' @importFrom scales pretty_breaks
 .plot_acc <- function(df, y_lim, optimal_ss, ...){
-    g <- ggplot(data = df, aes(x = sample))+
+    g <- ggplot(data = df, aes(x = as.factor(sample)))+
         geom_boxplot(aes(y = acc, group = sample, fill = fill_col), alpha = 0.5)+
         scale_fill_identity()+
         geom_point(aes(y = mean_acc))+
         geom_line(aes(y = mean_acc, group = 1), size = 0.75, color = "blue")+
         labs(x = "Simulated Sample Size Per Group", y = "Predictive Accuracy",
-             #title = sprintf("Classifier %s", alg),
-             subtitle = sprintf("Optimum accuracy achieved when sample size per group is : %s",
+             subtitle = sprintf("Optimal sample size per group is : %s",
                                 optimal_ss))+
         scale_y_continuous(breaks = scales::pretty_breaks(), limit = y_lim)+
-        scale_x_continuous(breaks = unique(df$sample))+
         theme_MSstats(...)+
         theme(plot.subtitle = element_text(face = "italic", color = "red"))
     
@@ -593,19 +592,24 @@ theme_MSstats <- function(x.axis.size = 10, y.axis.size = 10,
 #' @import data.table
[…]

### 2020-08-14 inst/sample_size_estimator/global.R

@@ -21,7 +21,6 @@ req_files <- list.files(src_dir, full.names = T)
 for (i in req_files) {
   tryCatch({
     source(i)
-    message(sprintf("Sourced file : %s", i))
   }, error = function(e){
     print(conditionMessage(e))
     print(sprintf("Error in sourcing : %s", i))
@@ -29,7 +28,7 @@ for (i in req_files) {
   
 }
 conn <<- .logGeneration()
-message("\n Log File saved at", conn$file)
+message("Log File saved at ", conn$file)
 
 #### GLOBAL VARS ####
 FORMATS_LIST <- list("Protein-level quantification" = "standard", 
@@ -102,20 +101,31 @@ tuning_params <- function(...){
 
 
 
-format_data <- function(format, count = NULL, annot = NULL, session = NULL, conn){
+format_data <- function(format, count = NULL, annot = NULL, transform = NULL,
+                        session = NULL, conn){
+  
   shiny::validate(shiny::need(format %in% FORMATS, "Undefined Format"))
   if(format == "standard"){
     .status(detail = "Importing Protein Abundance file", value = 0.4,
            session = session, log = conn$con)
     #read abundance data from the file path provided
-    wide <- fread(count$datapath, keepLeadingZeros = T)
+    wide <- read.csv(count$datapath, stringsAsFactors = FALSE)
     #No column names expected for the protein columns
-    setnames(wide, "V1", "Protein")
+    rownames(wide) <- wide[,1]
+     wide[,1] <- NULL
+    
+    if(transform != "No Transform"){
+      .status(detail = sprintf("Applying transformation of %s", transform),
+              value = 0.45, session = session, log = conn$con)
+      wide <- rapply(wide, f = get(transform), classes = c("numeric", "integer"),
+                     how = "replace")
+    }
+    
     name <- count$name
     .status(detail = "Importing Annotations file", value = 0.5,
            session = session)
     #read annotations from the file path provided
-    annot <- fread(annot$datapath, keepLeadingZeros = T)
+    annot <- read.csv(annot$datapath, stringsAsFactors = FALSE)
     name <- count$name
     
   }else if(format == "examples"){
@@ -147,7 +157,6 @@ format_data <- function(format, count = NULL, annot = NULL, session = NULL, conn
   .status(detail = "Creating Box Plots", value = 0.95, session = session,
          log = conn$con)
   .status(detail = sprintf("Dataset Name: %s", name), log = conn$con)
-  
[…]

### 2020-08-14 inst/sample_size_estimator/scripts/Methods.R

@@ -328,18 +328,19 @@
     consistent_cols <- !setequal(required.annotation, colnames(annotation))
     if(consistent_cols){
         packageStartupMessage(" Failure")
-        nf <- required.annotation[!colnames(annotation) %in% required.annotation]
+        nf <- required.annotation[!required.annotation %in% colnames(annotation)]
         stop("CALL_", func,"_",nf,
              " is not provided in Annotation, please check annotation file")
     }
+    anno_cols <- colnames(data)[!colnames(data) %in% 'Protein']
 
-    ic <- setequal(annotation$BioReplicate, colnames(data)) && 
-        nrow(annotation) == ncol(data)
+    ic <- setequal(annotation$BioReplicate, anno_cols) && 
+        nrow(annotation) == length(anno_cols)
     if(!ic){
         packageStartupMessage(" Failure")
         stop("CALL_",func,"_",
              "Please check the annotation file. 'Bioreplicate' must match with",
-             "the column names of 'data'.")
+             " the column names of 'data'.")
     }
 
     
@@ -591,19 +592,24 @@ theme_MSstats <- function(x.axis.size = 10, y.axis.size = 10,
 #' @import data.table
 qc_boxplot <- function(data = NULL ,annot = NULL){
     #create the interactive boxplot for all the different proteins found in the data
-    data <- as.data.table(data, keep.rownames = T)
-    setnames(data, 'rn', 'proteins')
-    data <- melt(data, id.vars = 'proteins', variable.name = "BioReplicate",
+    if(!'data.table' %in% class(data)){
+        data <- as.data.table(data, keep.rownames = T)
+        setnames(data, 'rn', 'Protein')
+    }
+    
+    data <- melt(data, id.vars = 'Protein', variable.name = "BioReplicate",
                  value.name = "Abundance")
     annot <- as.data.table(annot)
     data <- merge(data, annot, by = "BioReplicate") 
+    data[, Abundance := as.numeric(format(Abundance, digits = 2))]
     
     box_plot <- plotly::plot_ly(data = data[!is.na(Abundance)],
                                 y = ~Abundance, x = ~BioReplicate, color = ~Condition,
                                 type = "box") %>%
-        plotly::layout(xaxis = list(title="Biological Replicate",showticklabels = TRUE,
+        plotly::layout(xaxis = list(title = "Biological Replicate",
+                                    showticklabels = TRUE,
                                     tickangle = -45 ), 
-                       yaxis = list(title="Protein abundance"),
[…]

### 2020-08-14 inst/sample_size_estimator/server.R

@@ -27,6 +27,16 @@ function(session, input, output) {
   })
   ## Set maximum size of uploaded files to 300mb
   options(shiny.maxRequestSize = 300*1024^2)
+  #### Toggle control for transform
+  observeEvent(input$data_format,{
+    if(input$data_format == "examples"){
+      shinyjs::hide(id = "transform")
+    }else{
+      shinyjs::show(id = "transform")
+    }
+  })
+  
+  
   #### Toggle control for sidebar ####
   # Enable or disable fileInputs based on type of data selected
   observeEvent(input$data_format,{
@@ -40,7 +50,6 @@ function(session, input, output) {
     rv$seed <- -1
     rv$use_h2o <- F
     rv$classification <- NULL
-    
     disable_btns <- c("download_plots", "back_varimp", "fwd_varimp", 
                       "generate_report", "run_model","nav_to_exp",
                       "fwd","back","download_pca", "run_model")
@@ -53,7 +62,7 @@ function(session, input, output) {
       data <- .catch_faults(
         expr = format_data(format = input$data_format, count = input$standard_count,
                            annot = input$standard_annot, session = session,
-                           conn = conn),
+                           transform = input$transform, conn = conn),
         session = session, conn = conn
       )
       
@@ -448,7 +457,9 @@ function(session, input, output) {
                                                k = 10, session = session,
                                                conn = conn)
       }else{
-        stop("No h2o for you yet! Go Fish!")
+        shiny::showNotification("No h2o for you yet! Go Fish!", duration = NULL,
+                                type = "error")
+        validate(need(rv$use_h2o, "No h2o for you yet! Go Fish!"))
       }
       
     et <- Sys.time()
@@ -588,33 +599,39 @@ function(session, input, output) {
   output$generate_report <- downloadHandler(
     filename =  sprintf("Report_%s.pdf", format(Sys.time(), "%Y%m%d%H%M%S")),
     content = function(file){
-      tempReport <- file.path(tempdir(), "Report.Rmd")
-      file.copy("Report.Rmd", tempReport, overwrite = T)
-      
-      params <- list(data = rv$classification, use_h2o= rv$use_h2o,
-                     alg = names(MODELS)[which(MODELS %in% input$classifier)],
-                     sample = input$s_size,
-                     annot = ifelse(is.null(input$standard_annot$datapath),
-                                    "MSstatsSampleSize Package",
-                                    input$standard_annot$datapath),
-                     count = ifelse(is.null(input$standard_count$datapath),
[…]

### 2020-08-14 inst/sample_size_estimator/ui.R

@@ -78,6 +78,12 @@ dashboardPage(
                                             accept = EXTENSTIONS))
                          ),
                          fluidRow(
+                           column(width = 3,
+                                  selectInput("transform", "Base for Log Transform",
+                                              choices = c("No Transform","log","log10","log2"),
+                                              selected = "No Transform"))
+                         ),
+                         fluidRow(
                            column(width = 2,
                                   actionButton("import_data", "Import dataset", 
                                                icon = icon("file-import"),

### 2020-08-17 inst/sample_size_estimator/server.R
added ability to download logs

@@ -634,4 +634,11 @@ function(session, input, output) {
       })
     }
   )
-  }
+  #### Download Log button
+  output$log_down <- downloadHandler(
+    filename = "Log_MSstatsSSE.txt",
+    content <- function(file){
+      file.copy(conn$file, file)
+    }
+  )
+}

### 2020-08-17 inst/sample_size_estimator/ui.R

@@ -38,7 +38,13 @@ dashboardPage(
                tabName = "explore_simulated", icon = icon("project-diagram")),
       #### Analyze Simulation Tab ####
       menuItem("3: Plan experiment", 
-               tabName = "plan_experiment", icon = icon("vial"))
+               tabName = "plan_experiment", icon = icon("vial")),
+      br(),
+      br(),
+      br(),
+      downloadButton(outputId = "log_down", label = "Download Log",
+                     icon = icon("download"), 
+                     style = "margin-top: 25px; margin-left: 25px; margin-right: 25px; color: #000000",)
     )
   ),
   ##### Body ####

### 2020-08-25 inst/sample_size_estimator/global.R
change params to rds file

@@ -83,7 +83,7 @@ CSS_BUTTON_RUN <- "background-color: orange;
 #### Shiny functions ####
 tuning_params <- function(...){
   dots <- list(...)
-  params <- read.csv("params.csv", stringsAsFactors = F)
+  params <- readRDS("params.rds")
   op <- list("enable" = F, "ids" = params$id)
   
   if("Parameter Tuning" %in% dots$checkbox){
@@ -135,7 +135,6 @@ format_data <- function(format, count = NULL, annot = NULL, transform = NULL,
     wide <- MSstatsSampleSize::OV_SRM_train
     #examples data from the package
     annot <- MSstatsSampleSize::OV_SRM_train_annotation
-    #setnames(wide,"rn","Protein")
     name <- "Ovarian Cancer SRM study"
   }else{
     stop("Not Defined")

### 2020-08-25 inst/sample_size_estimator/server.R

@@ -1,8 +1,25 @@
 
 function(session, input, output) {
   # enables the helper functionality in the UI
-  shinyhelper::observe_helpers(help_dir = "help_mds")
   
+  wd <- getwd()
+  src_dir <- file.path(wd,"scripts")
+  req_files <- list.files(src_dir, full.names = T)
+  
+  for (i in req_files) {
+    tryCatch({
+      source(i)
+    }, error = function(e){
+      print(conditionMessage(e))
+      print(sprintf("Error in sourcing : %s", i))
+    })
+    
+  }
+  
+  conn <<- .logGeneration()
+  message("Log File saved at ", conn$file)
+  
+  shinyhelper::observe_helpers(help_dir = "help_mds")
   # stop the h2o cluster once the app is shutdown
   onStop(function() {
     try({

### 2020-08-25 R/run_estimator.R
update rd files

@@ -1,9 +1,0 @@
-#' @export
-MSstatsSSE_shiny <- function() {
-  appDir <- system.file("sample_size_estimator", package = "MSstatsSampleSize")
-  if (appDir == "") {
-    stop("Could not find example directory. Try re-installing `MSstatsSampleSize`.", call. = FALSE)
-  }
-  
-  shiny::runApp(appDir, display.mode = "normal")
-}
